---
title: "Validator Economics: Sync committee selection"
author:
- name: Sandra Johnson
  url: https://twitter.com/sandJohnson
  affiliation: Consensys Software Inc (Research Group)
date: "`r Sys.Date()`"
description: 
  Calculation and visualisation of sync committee selection probabilities for increased maximum effective balance.
editor_options: 
  markdown: 
    wrap: 72
---

```{r message=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(lubridate)
library(rmarkdown)
library(skimr)
library(tibble)
library(tidyverse)

options(digits=10)
options(scipen = 999) 

knitr::opts_chunk$set(dpi = if (knitr::is_latex_output()) 72 else 300,
                      echo=FALSE)
```
## Fetch R functions to use
```{r}
source("sync_functions.R")
```
## Sync committee selection

The probability of a validator being selected to participate in the sync committee
adheres to the following process: 
1) The validator indices are shuffled 
2) One candidate validator results from the shuffled index
3) A random byte is generated for this candidate (note this candidate may have been chosen before but more than likely with a different generated random byte)
4) A comparison is made between the candidate validator's effective balance * maximum bytes value (255) and MaxEB * random byte generated. If the first term is bigger than the second, then the validator is added to the sync committee.
5) Another iteration of the loop is executed until we have chosen enough validators (each member of the committee is present >= 1) to make up the sync committee



## PROCESS OF SELECTING SYNC COMMITTEE MEMBERS

 1. Shuffle index of validators (shuffled_index)
 2. Assign candidate index (candidate_index <- shuffled_index)
 3. Generate a random byte for that candidate
 4. Check if the candidate passes the test
    - compare the candidate validator's effective balance * maximum bytes value (255) and
      MaxEB * random byte generated. 
 5. Yes - a. add to sync committee  if membership count is not yet 512
        - b. increment membership counter
 6. No - do not increment membership counter & pass through the loop again

This is a similar process to proposer selection, but the difference here is that 
we are sampling WITH REPLACEMENT to select several validators to participate in a committee. Therefore each validator could be chosen more than once and the probabilities of being chosen do not change with each draw.
```{r}

#' EXAMPLE SCENARIO, PRE-EIP7521: 
#' ------------------------------
#' Total validator set size - 900,000 three stakers (A,B,C), each with 300,000 validators
#' Shuffle the entire set each time to draw the next candidate
# -------------------------------------------------------------------------------------
active_validators_df_now <- tibble(
  staker = sample(c(rep("A",300000), rep("B",300000), rep("C", 300000))),
  selected = rep(0,900000)
)

# Note we are assuming that all validators have 32 ETH EB
# Pre-EIP-7521 all validators will pass the sync committee check
# We are therefore only really looking at how many validators have been selected 
# more than once to join the sync committee

sync_committee_df_now <- assign_sync_committee(active_validators_df_now)
# What is the number of unique validators in the sync committee?
sync_committee_unique_df_now <- assign_unique_sync_mbrs(sync_committee_df_now) 
# As expected no duplicates when the active validator set is so large, 
# there were no validators elected to the sync committee more than once

# Let us also look at how the stakers fared. 
# Each staker had 300,000 validatotrs & 
# all validators had 32 ETH with MaxEB = 32 ETH (pre-EIP-7251)
sync_staker_totals_df_now <- assign_total_sync_stakers(sync_committee_df_now)


## TODO: Plot proportion of stake (each 1/3) against the proportional 
## ----  representation in the sync committee (A: 152/512, B: 187/512, C: 173/512)
##                                            (A: 0.298,   B: 0.365,   C: 0.338 )
##       Repeat with different seeds to get a distribution
# ======================================================================================
```
## Explore sync committee membership in post-EIP-7521 Ethereum
```{r}
## Sync committee selection post-EIP-7521
## =========================================================================================+
## Use current values of active validator set, ignoring staker groups (?), but              |
## distribution across various consolidated validators                                      |
## Validator set total = 900,000                                                            |
## Staker A: All single: 300,000 (=300,000)                                                 |
## Staker B: 1x: 2; 30x: 1; 64x: 4687 (=4,690)                                              |
## Staker C: 1x: 50,024; 2x: 25,000; 5x: 10,000; 10x: 5,000; 30x: 1,660; 64x: 784 (=92,468) |
## =========================================================================================+
active_validators_df_maxeb_now <- tibble(
  staker = c(rep("A",300000), rep("B",4690), rep("C", 92468)),
  EB = c(rep(32,300002), rep(960,1), rep(2048,4687), rep(32,50024), rep(64,25000), rep(160,10000), rep(320,5000), rep(960,1660),rep(2048,784)) 
)

# Shuffle the validator set
active_validators_df_maxeb_shuffled_now <- active_validators_df_maxeb_now[sample(1:nrow(active_validators_df_maxeb_now)), ] 

sync_committee_df_maxeb_now <- assign_maxeb_sync_committee(active_validators_df_maxeb_shuffled_now)
# Check for distinct validators. This run had 510 unique validators, so we had validators selected
# more than once (187738 & 380987 were selected twice - both are from staker B with 2,048 ETH each). They are
sync_committee_unique_df_maxeb_now <- assign_unique_sync_mbrs(sync_committee_df_maxeb_now)
sync_staker_totals_df_maxeb_now <- assign_total_sync_stakers(sync_committee_df_maxeb_now)
# Another interesting grouping would be the validator EB for each staker:
sync_staker_totals_EB_df_maxeb_now <- assign_total_sync_stakers_EB(sync_committee_df_maxeb_now)

# Total weight of sync committee (total staked ETH)
# =================================================
sync_committee_total_ETH_maxEB_now <- sum(sync_committee_df_maxeb_now$EB)  # 419,552 & 428704 
sync_committee_total_ETH_now <- 32 * 512      # 16,384



```
## Probabilities and Expected values
Using total active validator set of 900,000 (dataframes suffixed by _now)
```{r}
# Totals for each type of validator
active_validators_df_maxeb_now_EB_totals <- active_validators_df_maxeb_now %>%
  group_by(EB) %>%
  summarise(selected = n()) 

# Totals for each staker
active_validators_df_maxeb_now_staker_totals <- active_validators_df_maxeb_now %>%
  group_by(staker) %>%
  summarise(selected = n()) %>%
  mutate(prob = selected/397158)

# Totals by staker & validator type
# Add probability column, which is the probability of being selected as a 
# sync committee member
active_validators_df_maxeb_now_staker_EB_totals <- active_validators_df_maxeb_now %>%
  group_by(staker, EB) %>%
  summarise(total = n(), .groups = 'keep') %>%  # summarise with both grouping vars (default is 'drop_last')
  mutate(prob = total/397158) %>%
  arrange(EB,staker)

# Expected numbers in the sync committee
# Expected value = n*p 
expected_sync_mbrs <- active_validators_df_maxeb_now_staker_EB_totals %>%
  mutate(selectacceptprob = prob * case_when(
      EB == 32 ~ (1/64),
      EB == 64 ~ (2/64),
      EB == 160 ~ (5/64),
      EB == 320 ~ (10/64),
      EB == 960 ~ (30/64),
      EB == 2048 ~ 1,
      TRUE ~ 0)
      ) 

totalselectacceptprobs = sum(expected_sync_mbrs$selectacceptprob)

# Expected proportions by staker and EB (i.e. validator consolidation type)
expected_sync_mbrs <- expected_sync_mbrs %>%
  mutate(expected = (selectacceptprob/totalselectacceptprobs) * 512) %>%
  mutate(proportions = expected/total) 

# Check total selected = 512 
sync_committee_total <- sum(expected_sync_mbrs$expected)
```

